---
title: "Visualisation, </br>Shiny"
author: "Etienne Côme"
date: "9 Novembre 2023"
output:
  revealjs::revealjs_presentation:
    theme: white
    transition: none
    self_contained: true
    css: slides.css
  beamer_presentation:
    toc: false
    incremental: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Shiny
<link href="https://fonts.googleapis.com/css?family=Armata&display=swap" rel="stylesheet">

- Interactive interface / dashboard 
- Quite easy to implement
- All in R
- Debugging sometimes delicate because // javascript-html
- Requires a shiny server to be deployed 


## Shiny
### 2 components

- <span class="green">UI: describes the interface</span>
  - creation of the layout
  - buttons, sliders, plots, ... 
- <span class="red">Server: describes the application logic </span>
  - carries out the treatments 
  - defines how the interface updates when inputs change
  - makes the link between the inputs and the outputs

## <span class="green">UI</span>

```{r, eval=FALSE,echo=TRUE}
ui <- fluidPage(
    # Application title
    titlePanel("Old Faithful Geyser Data"),
    # Sidebar with a slider input for number of bins 
    sidebarLayout(
        sidebarPanel(
            sliderInput(inputId = "bins", label = "Number of bins:", 
                        min = 1, max = 50, value = 30)
        ),
        # Show a plot of the generated distribution
        mainPanel(plotOutput(outputId = "distPlot"))
    )
)
```

## <span class="green">UI</span>


### Different components :

- layout : page, layout, panel
- inputs : actionButton, dateInput, actionLink, sliderInput, radioButtons,...
- outputs : plotOutput, dataTableOutput, ... + leaflet

see la cheat sheet for details

### ! inputId, outputId
- id to reference input and output (defined in the ui)
- allow the server to access them

## <span class="red">Server</span>
```{r, eval=FALSE,echo=TRUE}
# fait le liens entre input et output
server <- function(input, output) {

    output$distPlot <- renderPlot({
        # generate bins based on input$bins from ui.R
        x    <- faithful[, 2]
        bins <- seq(min(x), max(x), length.out = input$bins + 1)

        # draw the histogram with the specified number of bins
        hist(x, breaks = bins, col = 'darkgray', border = 'white')
    })
}
```

## <span class="red">Server</span>

### Reactive part :
- renderPlot, renderImage,renderTable, ...

Th block will be re run each time one of the input it use change.

see cheat-sheet for details


## <span class="red">Server, advanced reactivity</span>
Structure the code to avoid duplicated treatments :
```{r, eval=FALSE,echo=TRUE}
data = reactive({})
data()
```

## <span class="red">Server, réactivité avancée</span>
Structure the code to avoid duplicated treatments :

- reactive : reactive block that return a reactive expression
- observe : reactive block without any output (side effect i/o)

```{r, eval=FALSE,echo=TRUE}
data = reactive({})
data()
renderPlot(data())
observe({
  input$saveButton 
  write.csv(data())  
})
```

## <span class="red">Server, advanced reactivity</span>
Isolate a code block :
```{r, eval=FALSE,echo=TRUE}
observe({
  input$saveButton  # Do take a dependency on input$saveButton

  # isolate a whole block
  data <- isolate({
    a <- input$valueA   # No dependency on input$valueA or input$valueB
    b <- input$valueB
    c(a=a, b=b)
  })
  writeToDatabase(data)
})
```


## <span class="red">Server, advanced reactivity</span>
Execute a block conditional to a specified list of inputs change
```{r, eval=FALSE,echo=TRUE}
data = eventReactive(input,{})
data()
observeEvent(input,{}})
```


## <span class="red">Server, advanced reactivity</span>
Deploy :

- shinyapp.io -> account creation
- button publish document
- get key and secret
- package the data needed by the app in an Rdata file in the app folder 
- use the load functio to load Rdata files

